---
title: "Language Statistics"
author: "Daniel Kasper"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---


# Disclaimer

In this task I chose to try and play around with the conversion of datatypes, how to webscrape and how to vizualise the data gathered. The task solved is not very pretty since it is a first attempt. With longer work-hours and more reading on GGplots syntax, better vizualisations could be made. Keep that in mind when looking at the results. 

## Loading the packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rvest)
library(dplyr)
library(tidyr)
library(stringr)
library(janitor)


```

## Using the webscraper on 2 different URLs


```{r url}
url <- "https://www.worldometers.info/world-population/population-by-country/"
# scrape the website
url_html <- read_html(url)


url2 <- "http://www2.harpercollege.edu/mhealy/g101ilec/intro/clt/cltclt/top100.html"
url_html2 <- read_html(url2)

```

## Converting the data to tables

```{r url-head}
whole_table <- url_html %>% 
 html_nodes("table") %>%
 html_table()  #str(whole_table) turns out to be a list
head(whole_table)



whole_table2 <- url_html2 %>%
	html_nodes("table") %>%
	html_table()
head(whole_table2)


```
## Converting tables to strings

```{r to table}
new_table <- do.call(cbind,unlist(whole_table, recursive = FALSE)) 
head(new_table) # ok, looks good, too bad it took 2 hours

new_table2 <- do.call(cbind,unlist(whole_table2, recursive = FALSE))
head(new_table2)

```


## Making a merged dataframe of the 2 scraped websites

Wrote

```{r join colummns from tables}

df <- as.data.frame(whole_table)
df2 <- as.data.frame(whole_table2)

joined_df <- merge(x=df, y=df2,by=NULL)

head(joined_df)

#write_csv(df,"data/population.csv")
#write_csv(df2,"data/language.csv")
 
```

## Converting chr strings and making vizualisations

Here I tried to properly convert the different datatypes which i had gathered so future vizualisations could benefit of this. I could have made a loop to ease the task but it was only 4 columns that needed to be converted.

Then I make a subset filter for checking which Languages are the most prominent in India by their population.


```{r language distribution in India}
library(ggplot2)
library(ggridges)

language_data <- read.csv("data/language.csv")

data <- language_data %>% 
	mutate(X4 = as.character(X4))  %>% 
	rename(Population = "X4") 

data2 <- data %>%
	mutate(X2 = as.character(X2)) %>%
	rename(LanguageName = "X2")

data3 <- data2 %>%
	mutate(X3 = as.character(X3)) %>%
	rename(PrimaryCountry = "X3")


data4 <- data3 %>%
	mutate(X1 = as.character(X1)) %>%
	rename(Rank = "X1")

language<- data4 %>%  filter(!row_number() %in% c(1))


languages_spoken <- language %>% 
  select(LanguageName, PrimaryCountry, Population, Rank) %>% 
  filter(PrimaryCountry == 'India') %>% 
  filter(!is.na(Population))

ggplot(data=languages_spoken, aes(x=Population,y=LanguageName)) + geom_col(color="blue", fill = "red") + theme(axis.text.x = element_text(angle = 90))

```

